---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
* Load the data
```{r}
activity_data <- read.csv("activity/activity.csv")
```

## What is mean total number of steps taken per day?
* Calculate the total number of steps taken per day
```{r}
step_sum_by_date <- aggregate(activity_data$steps, list(activity_data$date), sum, na.rm = TRUE)
names(step_sum_by_date) <- c("date", "steps")
step_sum_by_date_ignore_zero <- step_sum_by_date[which(step_sum_by_date$steps != 0), ]
print(step_sum_by_date_ignore_zero)
```

* Make a histogram of the total number of steps taken each day
```{r}
hist(step_sum_by_date_ignore_zero$steps, main = "steps per day", xlab = "steps")
```

* Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(step_sum_by_date_ignore_zero$steps)
median(step_sum_by_date_ignore_zero$steps)
```

## What is the average daily activity pattern?
* Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
five_minutes_interval_average <- tapply(activity_data$steps, activity_data$interval, mean, na.rm = TRUE)
plot(five_minutes_interval_average, type = "l")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(which.max(five_minutes_interval_average))
```

## Imputing missing values
* Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity_data))
```

* Create a new dataset that is equal to the original dataset but with the missing data filled in with the mean for that day.
```{r}
mean <- aggregate(activity_data$steps, list(activity_data$date), mean)
mean[which(is.na(mean[, 2])), 2] <- 0
for (row in which(is.na(activity_data$steps))) {
        activity_data[row, 1] <- mean[activity_data[row, 2], 2]
}
```

* Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.
```{r}
step_sum_by_date_fill_na <- aggregate(activity_data$steps, list(activity_data$date), sum)
names(step_sum_by_date_fill_na) <- c("date", "steps")
hist(step_sum_by_date_fill_na$steps, main = "steps per day", xlab = "steps")
mean(step_sum_by_date_fill_na$steps)
median(step_sum_by_date_fill_na$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
* Create a new factor variable in the dataset with two levels ¡V ¡§weekday¡¨ and ¡§weekend¡¨ indicating whether a given date is a weekday or weekend day.
```{r}
mean <- aggregate(activity_data$steps, list(activity_data$date), mean)
mean[which(is.na(mean[, 2])), 2] <- 0
for (row in which(is.na(activity_data$steps))) {
        activity_data[row, 1] <- mean[activity_data[row, 2], 2]
}
activity_data <- cbind(activity_data, weekdays(as.Date(activity_data$date)))
names(activity_data) <- c("steps", "date", "interval", "weekday")
wday <- vector()
for (row in 1 : nrow(activity_data)) {
	if (activity_data[row, 4] == levels(activity_data$weekday)[5] ||
		activity_data[row, 4] == levels(activity_data$weekday)[6]) {
		wday <- c(wday, "weekend")
	} else {
		wday <- c(wday, "weekday")
	}
}
activity_data$wday <- wday
```

* Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
options( warn = -1 )
activity_weekday <- data.frame()
activity_weekend <- data.frame()
for (row in 1: nrow(activity_data)) {
	if (activity_data[row, 5] == "weekday") {
		activity_weekday <- rbind(activity_weekday, activity_data[row, ])
	} else {
		activity_weekend <- rbind(activity_weekend, activity_data[row, ])
	}
}

agg_weekday <- aggregate(activity_weekday, list(activity_weekday$interval), mean, na.rm = TRUE)
agg_weekday$wday <- "weekday"
agg_weekend <- aggregate(activity_weekend, list(activity_weekend$interval), mean, na.rm = TRUE)
agg_weekend$wday <- "weekend"

library(lattice)
xyplot(steps ~ interval | wday, rbind(agg_weekday, agg_weekend), type = "l", layout = c(1, 2))
```